---
title: "607 Week 6 Assignment"
author: "Chirag Vithalani"
date: "March 4, 2016"
output: html_document
css: ./lab.css
---

<hr />

<ul class="a">
<li >

GOAL 1: Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. 

</li>
</ul>

<ul class="b">
<li>
tidyr provides three main functions for tidying your messy data: gather(), separate() and spread().
</li>
</ul>

<ul class="c">
<li>
Reading CSV file and printing data

```{r}

library(tidyr)
flightData<-read.csv("https://raw.githubusercontent.com/chirag-vithlani/607/master/week6/week6.csv", header= TRUE,sep=",",na.strings = "?",stringsAsFactors=FALSE)
flightData
```
</li>

<li>
Missing headers are names as X and X.1, so rename headers
```{r}
names(flightData)[names(flightData) == "X"] <- "Airline"
names(flightData)[names(flightData) == "X.1"] <- "Arrival"

```
</li>

<li>
Repeating airline name where it is missing ( this is hard coded, but this step should be done programmatically )
```{r}
flightData[2, 1] <- "ALASKA"
flightData[4, 1] <- "AM West"
flightData

```
</li>
<li>
Using `gather()` function.
`gather()` takes multiple columns, and gathers them into key-value pairs: it makes "wide" data longer.
```{r}
tidy <- gather(flightData, "City", "Count", 3:7) 
head(tidy)
```
</li>

<li>

Using `spread()` function.
`spread()` takes two columns (a key-value pair) and spreads them in to multiple columns, making "long" data wider

```{r}

#tidy$Count <- as.numeric(gsub(",", "", tidy$Count))
tidy <- spread(tidy, "Arrival", Count)
tidy


```
</li>
</ul>

<ul class="b">
<li>

dplyr implements the following verbs useful for data manipulation:
<pre>
select(): focus on a subset of variables
filter(): focus on a subset of rows
mutate(): add new columns
summarise(): reduce each group to a smaller number of summary statistics
arrange(): re-order the rows
</pre>

</li>
</ul>

<ul class="c">

Choosing columns: select
```{r}
suppressMessages(library(dplyr))
head(select(tidy, Airline))
```
or use minus to hide that column
```{r}
head(select(tidy, -Airline))
```
</li>

<li>
Using filter

```{r}
filter(tidy,Delayed==62)
```
</li>


<li>
Using mutate : Add new variables

```{r}
tidy=mutate(tidy,Total = Delayed + `On Time`)
head(tidy)
```
</li>

<li>
Using summarise : Reduce variables to values

```{r}
tidy <- mutate(tidy, Total = Delayed + `On Time`, PercentDelayed = Delayed / Total * 100)
tidy <- arrange(tidy, City, PercentDelayed)
tidy
```
</li>
</ul>
<ul class="a">
<li>
GOAL 2: Perform analysis to compare the arrival delays for the two airlines.
</li>

</ul>


<ul class="c">
<li>

 As we can see Alaska airline is delayed by 11% and AM West delayed by 18%.

```{r}
delays <- tidy %>% group_by(Airline) %>% summarise(MeanPercent = round(mean(PercentDelayed), 0))
delays
```
</li>
</ul>