---
title: "607_Week2Assignment"
author: "Chirag Vithalani"
date: "January 30, 2016"
output: html_document
---

**607_Week2Assignment**

Your task is to choose one dataset, then study the data and its associated description of the data (i.e. "data dictionary"). You should take the data, and create an R data frame with a subset of the columns (and if you like rows) in the dataset.Your deliverable is the R code to perform these transformation tasks.

**Reading bridge data** 

```{r}
bridgev1file<-read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/bridges/bridges.data.version2", header= FALSE, sep=",",na.strings = "?")
head (bridgev1file)
```
**Read data dictionary** (simplied)

```{r}
bridge_dictionay <- read.table("./data_dictionary.txt",row.names = 1, sep=":")
print(bridge_dictionay)

```
**Apply headers to data by reading from data_distionary**
```{r}
names(bridgev1file) <- row.names(bridge_dictionay)
head(bridgev1file)
```
**I expected all 'HIGHWAYS' with 'LONG' length are made up of 'IRON' or 'STEEL',but there are few HIGHWAYS made up of 'WOOD'**

```{r}

bridge_sub_df <- subset(bridgev1file, select=c(5,6,7,10))
print(bridge_sub_df)
```
