---
title: "607 Week 4 Assignment"
author: "Chirag Vithalani"
date: "February 16, 2016"
output: html_document
css: ./lab.css
---



<hr />

<ul class="a">
<li >
3. Copy the introductory example. The vector name stores the extracted names.
<pre>
R> name
[1] "Moe Szyslak" "Burns, C. Montgomery" "Rev. Timothy Lovejoy"
[4] "Ned Flanders" "Simpson, Homer" "Dr. Julius Hibbert"
</pre>




<ul class="b">
<li >
(a)  Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.


Below is what we have

```{r}
library(stringr)
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555
-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```
<ul class="c">
<li>

To get standard names below are steps

<ol>
<li>we need to convert list to dataframe</li>
<li>rename it</li>


```{r}
#convert list to dataframe
namesdf<-do.call(rbind, lapply(name, data.frame, stringsAsFactors=FALSE))
#rename column
namesdf$names<-namesdf$X..i..
```
<li>If name contains comma, that means last name is first and first name is last.so re-arrange it</li>


```{r}
namesdf$stdFormatNames<-ifelse(grepl( ",",namesdf$names),paste(word(namesdf$names,-1),word(namesdf$names,1)),namesdf$names)
```

<li>remove prefixes and commas</li>
  
```{r}
namesdf$stdFormatNames<-gsub("Rev.|Dr.|,","", namesdf$stdFormatNames)
namesdf$stdFormatNames
```
</ol>
</li>
</ul>

</li>
</ul>

<ul class="b">
<li>

(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).
</li>
</ul>

<ul class="c">
<li>
Using str_detect to check if there is title in name

```{r}
namesdf$names

namesdf$hasTitle<-str_detect(namesdf$names, "Rev.|Dr.")
namesdf$names
namesdf[,c("names","hasTitle")]
```
</li>
</ul>

<ul class="b">
<li>
(C) Construct a logical vector indicating whether a character has a second name.
</li>
</ul>

<ul class="c">
<li>

Spliting by space

```{r}
namesdf$stdFormatNames
grepl( " ",str_trim(namesdf$stdFormatNames))
```

</li>
</ul>
</li>
</ul>

<ul class="a">
<li>
7. Consider the string <title>+++BREAKING NEWS+++</title>. We would like to extract the first HTML tag. To do so we write the regular expression <.+>. Explain why this fails and correct the expression.


<ul class="c">
<li>
R applies greedy quantification. This means that the program tries to extract the greatest Greedy quantification and how to avoid it possible sequence of the preceding character. As the . matches any character, the function returns the greatest possible sequence of any characters before a sequence of sentence. 


```{r}

stringToSearch<-"<title>+++BREAKING NEWS+++</title>"
str_extract(stringToSearch,"<.+>")

```

We can change this behavior by adding a ? to the expression in order to signal that we are only looking for the shortest possible sequence of any characters before a sequence of sentence.


```{r}
str_extract(stringToSearch, "<.+?>")
```
</li>
</ul>

</li>
</ul>

<ul class="a">
<li>
8. Consider the string (5-3)^2=5^2-2*5*3+3^2 conforms to the binomial theorem. We would like to extract the formula in the string. To do so we write the regular expression [^0-9=+*()]+. Explain why this fails and correct the expression.

<ul class="c">
<li>
Putting the caret at the beginning of a character class does the inverse.
keeping it at the end resolves the problem.

```{r}
binomial_str <- "(5-3)^2=5^2-2*5*3+3^2 conforms to the binomial theorem."
str_extract(binomial_str, "[^0-9=+*()]+")
str_extract(binomial_str, "[0-9=+*()^-]+")

```
</li>
</ul>

</li>
</ul>