---
title: "Data Acquisition and Management  - Project 2"
author: "Chirag Vithalani"
date: "March 12, 2016"
output: 
  html_document: 
    highlight: pygments
    theme: cerulean
    toc: false
    number_sections: false
---

## Fun with world bank data {.tabset .tabset-fade .tabset-pills}


#### Here goal is to read "wide" datasets and and use tidyr and dplyr as needed to tidy and transform data.
<hr />
### Data set 1 : World Population 
<hr />

 

```{r echo=FALSE,message=FALSE}


suppressWarnings(library(tidyr))  # for transformation
suppressWarnings(library(ggplot2)) #to draw graphs


```
####Step 1 : Reading data file

```{r  message=FALSE}
world_population<-read.csv("https://raw.githubusercontent.com/chirag-vithlani/607/master/project2/world_populationOnlyCountry.csv", header= TRUE,sep=",",na.strings = "?",stringsAsFactors=FALSE)

```

####Step 2 : Selecting only required columns using `dplyr`'s select() function
i.e. selecting only year last 30 years of data ( with gap of 10 years )

```{r  message=FALSE}

library(dplyr)
# World population subset
wp_subset<- select(world_population,Country.Name,Country.Code,X1993,X2003,X2013)
head(wp_subset,1)
#other way to achieve same thing
#wp_subset<-subset(world_population, select=c("Country.Name","X1993","X2003","X2013"))

```
#### Step 3 : Rename columns using `rename()` function of plyr

```{r message=FALSE}

suppressWarnings(detach("package:dplyr", unload=TRUE)) # this is required step as dplyr and plyr doesn't work togather

# Getting reuired world population data
suppressWarnings(library(plyr))
wp<-rename(wp_subset, c("Country.Name"="country","Country.Code"="code","X1993"="1993", "X2003"="2003","X2013"="2013"))

head(wp,1)

```

#### Step 4 : Using `gather()` function to get data year wise.

Putting 'year' information in rows

```{r message=FALSE}
suppressWarnings(library(dplyr)) #dplyr is a powerful R-package to transform and summarize tabular data with rows and columns.

# gather() makes wide data longer (here we are removing 3rd to 5th column)
wp_year_wise<-gather(wp, "Year", "Count", 3:5)

head(wp_year_wise,1)

```
#### Step 5 : Remove rows which doesn't contain data using `filter` function.

```{r message=FALSE}

suppressWarnings(library(plotly))

#wp_subset$percent<-round((wp_subset$X2013/total_wp13)*100,2)
wp_year_wise <- wp_year_wise %>% filter(!is.na(wp_year_wise$Count))

```
#### Step 6 : Now we need to get percentage of country population with resect to world.

This is bit complex step,
although formula is simple, just divide country population with total world population, but we need to find this total population with respect to ( in that particular ) year.

First finding total population in ( that !! ) year.

we do this by using functions `mutate()`, `sum()` and `group_by()`

<i>(Note : There was row in source data for 'world` too (that would make life easier) , but I choose to remove it and kept only country specific data.)</i>

```{r message=FALSE}
wp_year_wise<- wp_year_wise %>% 
    group_by(Year) %>% 
    mutate(Total_Population=sum(as.numeric(Count)))

#just to see what data generated (after above step)
subset( wp_year_wise, wp_year_wise$code=='USA')
```
#### Step 7 : Now calculating percentage population.

```{r}

wp_year_wise$percent<-round((wp_year_wise$Count/wp_year_wise$Total_Population)*100,4)
#save(wp_year_wise, file="mydata.RData")

```
#### Step 8 : Now is the time for fun ;)

We all know China and India are most populous countries,but how much with respect to world population.

```{r message=FALSE}

Population_India_China_2013<-subset(wp_year_wise,wp_year_wise$Year==2013 & (wp_year_wise$code=='IND' | wp_year_wise$code=='CHN'))

India_China_2013_Sum_percent<-sum(Population_India_China_2013$percent)

dataset<-data.frame(labels = c(" India + China "," Rest"),values = c(India_China_2013_Sum_percent,100-India_China_2013_Sum_percent))

plot_ly(dataset, labels = labels, values = values, type = "pie") %>% 
    layout(title = "China + India Poulation v/s Rest Of Wolrd")


```



#### Step 8 : Now choosing US,China & India specific data.

```{r message=FALSE}

#library(sqldf)
wpUCI<-subset(wp_year_wise,wp_year_wise$country=='India' | wp_year_wise$country=='China' |  wp_year_wise$country=='United States')
head(wpUCI,1)

```
#### Last Step : Conclusion with bubble plot
##### As we can see in below plot, how countries population changed over time. 
##### And this also shows relative population of country with respect to 
##### world population ( when we mouse hover on bubble, and color of bubble )
##### popluation of other coutries ( by observing the size of bubble )


```{r message=FALSE}
plot_ly(wpUCI, x = Year, y = code, size = percent,color = percent,opacity = percent, mode = "markers")



```

US population is comparatively small, we always knew this, but this gives nice relative insight.Like initially China was leading, but now India is caching up. US population hasn't changed much ( with respect to world ) as color of bubble doesn't change with time.



<hr />
###Data set 2 : World Forests 
##### This dataset is almost similar to population one, so doing almost same operations here,  but plotting it differently.
<hr />

#### Step 1 : Read forests data.
```{r message=FALSE}
world_forests<-read.csv("https://raw.githubusercontent.com/chirag-vithlani/607/master/project2/forests.csv", header= TRUE,sep=",",na.strings = "?",stringsAsFactors=FALSE)

```

#### Step 2 : Selecting only required columns using `dplyr`'s select() function
i.e. selecting only year last 30 years of data ( with gap of 10 years )

```{r message=FALSE}
wf_subset<-subset(world_forests, select=c("Country.Name","Country.Code","X1993","X2003","X2013"))
head(wf_subset)

#need to detach dplyr as plyr and dplyr both doesn't work 
detach("package:dplyr", unload=TRUE)

```
#### Step 3 : Rename the columns

```{r message=FALSE}
suppressWarnings(library(plyr))
wf<-rename(wf_subset, c("Country.Name"="country","Country.Code"="code","X1993"="1993", "X2003"="2003","X2013"="2013"))
head(wf,1)

library(dplyr)

wf_year_wise<-gather(wf, "Year", "Count", 3:5)
head(wf_year_wise,1)
wf <- wf_year_wise %>% filter(!is.na(wf_year_wise$Count))
head(wf,1)


```


#### Step 4 : Now what we need is "Total Forest Area of world-in that particular year". Using `mutate()` and `group_by()` to achieve this.

```{r message=FALSE}

wf<- wf %>% 
    group_by(Year) %>% 
    mutate(Total_Forest_Area=sum(as.numeric(Count)))

```

#### Step 5 : Calculating percentage of forest land with respect to total forest land in world - in that particular year.

```{r message=FALSE}

wf$percent<-round((wf$Count/wf$Total_Forest_Area)*100,4)

```

####Plotting on world map

```{r}
#df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv')
l <- list(color = toRGB("grey"), width = 0.5)
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

plot_ly(wf, z = Count, text = code, locations = code, type = 'choropleth',        color = Count, colors = 'Greens', marker = list(line = l),colorbar = list(ticksuffix = '%', title = 'Forests Color Map')) %>% layout(title = 'Forests',geo = g)

```

I didn't know Finaland has huge forestry.

### Merging two data sets 
#####( not very useful info, but shows use of functions )


Three columns are common among two datasets ( world population & forest )

```{r message=FALSE}

Population_And_Forests<-merge(wf,wp_year_wise,by=c("country","code","Year"))
head(Population_And_Forests,1)

```
####Forest land with respect to population -- i.e forest land percentage per person by country

```{r message=FALSE}

options(scipen = 999)
Population_And_Forests$forests_by_population<-round(((Population_And_Forests$Total_Forest_Area/Population_And_Forests$Total_Population)*100),8)



```
####Using `arrage()` function to sort.

```{r message=FALSE}

head(Population_And_Forests)
sorted<-arrange(Population_And_Forests,-forests_by_population)
head(sorted)
subset(sorted,select=c("country","forests_by_population"),sorted$code=="IND"|sorted$code=="USA"|sorted$code=="CHN")

```

