---
title: "607_Assignment2"
author: "Chirag Vithalani"
date: "February 13, 2016"
output: html_document
---

###Your task is to create an R dataframe that shows rates of tuberculosis infection by country. You should use the information from the tb database and the attached population.csv file. Your R dataframe should have the following columns:Country, Year, Rate. Where Rate is defined as Cases/Population.

####Step 1: So we need Country,Year and Rate. Getting first two column is easy,so let us do that first.

```{r}
population<-read.csv("https://raw.githubusercontent.com/chirag-vithlani/607/master/week3/population.csv", sep=",")
head(population)
```

So we have first two columns.
to calculate rate, we need cases.

####Step 2: We need to get cases from database.
Use below query to get data to R, once data is uploaded to database using SQL.

SELECT YEAR,COUNTRY,SUM(MALECASES)
FROM 
(
  
  SELECT  TB1.*,TB1.CHILD+TB1.ADULT+TB1.ELDERLY AS maleCases  
  FROM TB TB1 WHERE TB1.SEX='MALE' 
  
  UNION 
  
  SELECT  TB2.*,TB2.CHILD+TB2.ADULT+TB2.ELDERLY AS femaleCases 
  FROM TB TB2 
  WHERE TB2.SEX='FEMALE' 
  
) MASTER
GROUP BY YEAR,COUNTRY

```{r, echo=FALSE}

library(RMySQL)
mydb = dbConnect(MySQL(), user='root', password='Master31', dbname='tb', host='localhost')

```

```{r}
rs = dbSendQuery(mydb, "select year,country,sum(malecases) as total  from  (select  tb1.*,tb1.child+tb1.adult+tb1.elderly as malecases  from tb tb1 where tb1.sex='male' union select tb2.*,tb2.child+tb2.adult+tb2.elderly as femalecases from tb tb2 where tb2.sex='female' ) master group by year,country")

data = fetch(rs, n=-1)
subset(data,data$country=='Afghanistan')
```

####Step 3 :Now join two frames with joining key being Year and Country.

```{r}
library(dplyr)
fullData=inner_join(data,population)
fullData$rate=fullData$total/fullData$population
finalFrame=subset(fullData,select=c(1,2,5))
head(finalFrame)
```


